My title

### 2022-02-04 outputs/paper/paper.Rmd
commiting work

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Arsh Lakhanpal
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
library(opendatatoronto)
library(dplyr)
library(tidygeocoder)
library(lubridate)
library(janitor)
library(ggmap)
```

@@

```{r}
#covid_cases <- COVID19_cases
covid_cases <- clean_names(covid_cases)
covid_cases <- subset(covid_cases, select = -c(1,2,7,8))
```
The statistical programming language used to work with the data in this paper is R (add citation) along with packages like tidyverse, knitr. The "COVID-19 Cases in Toronto" dataset was obtained from Open Data Toronto which houses a broad number of datasets. This data set was published by Toronto Public Health and for this paper, this dataset was officially imported from the opendatatoronto package in R.

@@

@@

covid_cases <- covid_cases %>% 
  mutate(month_of_diagnosis = month(episode_date,
                  label = TRUE,
                  abbr = FALSE))

@@

covid_cases2020 %>% ggplot(mapping = aes(x= month_of_diagnosis)) + geom_bar()
covid_cases2021 %>% ggplot(mapping = aes(x= month_of_diagnosis)) + geom_bar()
"This is used to see which month had the most cases, see where it lines up with lockdowns"

@@

#covid_cases %>% count(month_of_diagnosis)
#covid_cases %>% count(outcome)
#covid_cases %>% count(source_of_infection)

@@

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Arsh Lakhanpal
thanks: "Code and data are available at: 
https://github.com/lakhan99/starter_folder-main"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The city of Toronto has been able to collect data pertraining to COVID-19 cases demographic and geographic variables since January 2020. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

library(patchwork)
library(tidyverse)
library(palmerpenguins)
library(opendatatoronto)
library(dplyr)
library(tidygeocoder)
library(lubridate)
library(janitor)
library(ggmap)
covid_cases <- read_csv(here::here("inputs/data/covid_cases.csv"))

@@

```{r}
covid_cases <- clean_names(covid_cases)
#covid_cases <- subset(covid_cases, select = -c(1,2,7,8))
```

  The statistical programming language used to work with the data in this paper is R (add citation) along with packages like tidyverse, knitr. The "COVID-19 Cases in Toronto" dataset was obtained from Open Data Toronto which houses a broad number of datasets. This data set was published by Toronto Public Health and for this paper, this dataset was officially imported from the opendatatoronto package in R.

@@

COVID-19, like many other viruses thrives in colder drier climates and Figure \@ref(fig:figone) . To do this, two new variables named "month_of_ diagnosis" and "year_of_diagnosis" were mutated using the tools from the dplyr package. The following two graphs show the number of covid cases for both 2020 and 2021 as those are the only two complete years that we have have to live with the virus.

```{r figone, echo=FALSE}

covid_cases <- covid_cases %>% 
  mutate(month_of_diagnosis = month(episode_date,
                  label = TRUE,
                  abbr = TRUE))

fig2020 <- covid_cases2020 %>% ggplot(mapping = aes(x= month_of_diagnosis)) + geom_bar() + labs(x = "Month of 2020", y = "Number of COVID cases")

fig2021 <- covid_cases2021 %>% ggplot(mapping = aes(x= month_of_diagnosis)) + geom_bar() + labs(x = "Months of 2021", y = "Number of COVID cases")

monthly_figs <- (fig2020 + fig2021)
monthly_figs + plot_annotation(caption = "Figure 1: COVID-19 Cases for every month in 2020 and 2021")
```
```{r tableone, echo=FALSE}

D <- count(covid_cases2020 %>% 
  filter(month_of_diagnosis == "Dec", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

S <- count(covid_cases2020 %>% 
  filter(month_of_diagnosis == "Sep", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

N <- count(covid_cases2020 %>% 
  filter(month_of_diagnosis == "Nov", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

D1 <- count(covid_cases2021 %>% 
  filter(month_of_diagnosis == "Dec", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

S1 <- count(covid_cases2021 %>% 
  filter(month_of_diagnosis == "Sep", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

N1 <- count(covid_cases2021 %>% 
  filter(month_of_diagnosis == "Nov", source_of_infection == "Close Contact" | source_of_infection == "Community" | source_of_infection == "Household Contact"))

September <- c(3012,2806)
November <- c(5029, 1932)
December <- c(9047, 3975)

holiday_case <- data.frame(September, November, December)
rownames(holiday_case) <- c("2020","2021")

holiday_case<- knitr::kable(holiday_case, caption = "Table 1: Number of covid cases from contact during holiday months")
holiday_case
```

In 2020, we can see how the number of covid cases were highest during the fall and winter months and how that trend carries onto early 2021 which should be what is expected. The number of cases tend to increase near the winter 2020 and that trend continued in until April 2021 where the cases began to continuously decrease. One thing to notice is that in both years, there are a a higher amount  of COVID cases comparatively speaking to other months. One of the possible reasons explaining this rise in cases during this time is the fact that this is around the holidays when families tend to get together. Table \@ref(tab:tableone) shows the count for COVID-19 cases that were transmitted through contact which includes close contact, household contact and community contact. For both years, we notice that the month of December had the highest number of COVID-19 cases, likely because of Christmas. Another thing to consider is that in December of 2020, there was a new and more contagious variant of the COVID-19 virus called the delta variant. Similarly, at the end of November 2021, we got the omicron variant, one that was even more contagions than the delta variant which also explains the high case counts in December of 2020 and 2021.

shows us  
(talk about possible lockdown/vaccine trends)

We can further focus on the covid cases for every month and which age group seemed to be most affected by covid throughout this pandemic.

@@

@@

#seperate this into 2020,2021 as well?

```
These bar charts..

@@

#covid_cases %>% count(month_of_diagnosis)
#covid_cases %>% count(outcome)
#propo <- covid_cases %>% count(source_of_infection)
#type <- unique(covid_cases[c("source_of_infection")])

#prop.table(propo)

@@

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
#ggplot(penguins, aes(x = island, fill = species)) +
 # geom_bar(alpha = 0.8) +
  #scale_fill_manual(values = c("darkorange","purple","cyan4"),
   #                 guide = FALSE) +
  #theme_minimal() +
  #facet_wrap(~species, ncol = 1) +
  #coord_flip()
```
